---
title: "Tractor Burden Test Simulation"
author: "Taotao Tan"
date: "12/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Here we are simulation a dataset to test Tractor burden test method

pre-defined parameters
```{r}
# sample size
num_sample = 5000
num_vars = 100

# simulate global admixture proportion with a beta distribution, E(admix) = 0.8
admix = rbeta(num_sample, 8, 2)

# minor allele frequency
AFR_maf = 0.05
EUR_maf = 0.05

# how many risk alleles are responsible for the phenotype
num_risk_alleles = 3

# effect size for AFR, EUR, and global admixture 
AFR_eff_size = 0
EUR_eff_size = 2
admix_eff_size = 1.2

# error term size
error_eff_size = 2


plot(density(admix), main = "distribution of percentage of admixture")
```


We assume the population is two way admixed with 80% AFR and 20% EUR in average

This could be modeled with a beta distribution $\text{Beta}(\alpha = 8, \beta = 2)$. The expected proportion of AFR compoents are 80%. (Is that a reasonable distribution to draw from?)

```{r}
# generate AFR specific hap count 
AFR_hapcount = data.frame(t(sapply(admix, function(x){
  return(rbinom(num_vars, size = 2, prob = x))
})))

# generate EUR specific hap count
EUR_hapcount = 2 - AFR_hapcount
```



Generate dosage of risk alleles based on: hap count, ancestry-specific probability of having risk alleles

```{r}
Generate_dosage = function(sample, ancestry){
  return(sapply(sample, function(x){return(rbinom(1, size = x, prob = ancestry))}))
}

# generate AFR risk allele dosage
AFR_dosage = data.frame(t(apply(AFR_hapcount,1, Generate_dosage, AFR_maf)))

# generate EUR risk allele dosage
EUR_dosage = data.frame(t(apply(EUR_hapcount,1, Generate_dosage, EUR_maf)))

# generate TOT (total) copies of risk allele dosage
TOT_dosage = AFR_dosage + EUR_dosage




hist(rowSums(AFR_dosage), main = "distribution of total AFR risk allele dosage")
hist(rowSums(EUR_dosage), main = "distribution of total EUR risk allele dosage")
hist(rowSums(TOT_dosage), main = "distribution of total risk allele dosage")



```




```{r}
risk_alleles = sample(num_vars, size =num_risk_alleles, replace = F)
admix_linear = log(admix/(1-admix))  # is this transformation necessary?
error_term = rnorm(num_sample, sd = error_eff_size)


y = rowSums(AFR_dosage[,risk_alleles] * AFR_eff_size) + rowSums(EUR_dosage[,risk_alleles] * EUR_eff_size) + admix_linear * admix_eff_size + error_term
```






Modeling
```{r}
# library(plotly)
df = data.frame(cbind(rowSums(AFR_dosage), rowSums(EUR_dosage), rowSums(TOT_dosage), admix, y))
colnames(df) = c("AFR", "EUR", "TOT", "admix", "pheno")
# plot_ly(df, x = ~AFR, y = ~EUR, z = ~pheno, color = ~admix, colors = c('#BF382A', '#0C4B8E'))


```






```{r}
df$admix_linear = log(admix/(1-admix))

summary(lm(pheno ~ AFR + EUR + admix_linear, data = df))$coefficients[2:3,4]




```






```{r}
library(tidyr)
Generate_dosage = function(sample, ancestry){
  return(sapply(sample, function(x){return(rbinom(1, size = x, prob = ancestry))}))
}


SIMULATOR <- function(param){ 
    num_sample = param[1]
    num_vars= param[2]
    admix_param1= param[3]
    admix_param2= param[4]
    AFR_maf= param[5]
    EUR_maf= param[6]
    num_risk_alleles= param[7]
    AFR_eff_size= param[8]
    EUR_eff_size= param[9]
    admix_eff_size= param[10]
    error_eff_size= param[11]
  
    admix = rbeta(num_sample, admix_param1, admix_param2)
    
    # AFR hapcount, generated from admix proportion with a binomial distribution (n = 2, prob = admix)
    AFR_hapcount = data.frame(t(sapply(admix, function(x){ return(rbinom(num_vars, size = 2, prob = x))})))
    EUR_hapcount = 2 - AFR_hapcount
    
    # generate AFR risk allele dosage
    AFR_dosage = data.frame(t(apply(AFR_hapcount,1, Generate_dosage, AFR_maf)))

    # generate EUR risk allele dosage
    EUR_dosage = data.frame(t(apply(EUR_hapcount,1, Generate_dosage, EUR_maf)))

    # generate TOT (total) copies of risk allele dosage
    TOT_dosage = AFR_dosage + EUR_dosage
    
    
    # determine which variants are risk alleles
    risk_alleles = sample(num_vars, size =num_risk_alleles, replace = F)
    admix_linear = log(admix/(1-admix))  # is this transformation necessary?
    error_term = rnorm(num_sample, sd = error_eff_size)
    

    # the outcome is a function of true risk alleles from EUR and AFR,
    # global ancestry (transformed into linear way), and an error term
    y = rowSums(AFR_dosage[,risk_alleles,drop = FALSE] * AFR_eff_size) + rowSums(EUR_dosage[,risk_alleles,drop = FALSE] * EUR_eff_size) + admix_linear * admix_eff_size + error_term
    
    
    # sum over AFR and EUR risk alleles, and perform a linear regression
    df = data.frame(cbind(rowSums(AFR_dosage), rowSums(EUR_dosage), rowSums(TOT_dosage), admix_linear, y))
    
    colnames(df) = c("AFR", "EUR", "TOT", "admix", "pheno")
    
    summary(lm(pheno ~ AFR + EUR + admix_linear, data = df))$coefficients[,4] 
    
    return(c(
      TOT = summary(lm(pheno~TOT + admix_linear, data = df))$coefficients[2,4],
      summary(lm(pheno ~ AFR + EUR + admix_linear, data = df))$coefficients[2:3,4]
    ))                     
}


params = crossing(num_sample = c(1000, 5000, 10000),
         num_vars = 100, 
         admix_param1 = 8, 
         admix_param2 = 2,
         AFR_maf = c(0.01, 0.05), 
         EUR_maf = c(0.01, 0.05), 
         num_risk_alleles = c(1, 5, 20),
         AFR_eff_size = c(0, 1, 2), 
         EUR_eff_size = c(0, 1, 2), 
         admix_eff_size = 1,
         error_eff_size = 2)



res = t(apply(params[40:50,], 1, SIMULATOR))

write.table(res, row.names = F, col.names = T, quote = F)


```




