---
title: "Vis_simulation"
author: "Taotao Tan"
date: "12/23/2021"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```



```{r, echo=FALSE}
files = paste0("Simulation_res/res_", 1:100)
all_res = lapply(files, read.csv, sep = " ")

```


```{r, echo=FALSE}
library(tidyr)

params = crossing(num_sample = c(1000, 5000, 10000),
         num_vars = 100, 
         admix_param1 = 8, 
         admix_param2 = 2,
         AFR_maf = c(0.01, 0.05), 
         EUR_maf = c(0.01, 0.05), 
         num_risk_alleles = c(1, 5, 20),
         AFR_eff_size = c(0, 1, 2), 
         EUR_eff_size = c(0, 1, 2), 
         admix_eff_size = 1,
         error_eff_size = 2)


# write.table(params, "parameters.csv", quote = F, sep = "\t", row.names = F, col.names = T)
```

```{r, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)

Retrieve_data <- function(num_sample, AFR_maf, EUR_maf, num_risk_alleles, AFR_eff_size, EUR_eff_size){
  boo = (params[,"num_sample"] == num_sample) &
    (params[,"AFR_maf"] == AFR_maf) &
    (params[,"EUR_maf"] == EUR_maf) &
    (params[,"num_risk_alleles"] == num_risk_alleles) &
    (params[,"AFR_eff_size"] == AFR_eff_size) &
    (params[,"EUR_eff_size"] == EUR_eff_size)
  
  lst_df = lapply(all_res, function(x){return(x[boo,])})
  res = bind_rows(lst_df, .id = "Simulation_num")
  return(res)
}

```



# Different sample size

#### The following conditions remain unchanged:
AFR_maf = 0.05  

EUR_maf = 0.05  

num_risk_alleles = 5  

AFR_eff_size = 1  

EUR_eff_size = 1

```{r, echo=FALSE}
test1 = Retrieve_data(num_sample = 1000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 1)
test2 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 1)
test3 = Retrieve_data(num_sample = 10000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 1)

test_melt1 = melt(test1)
test_melt2 = melt(test2)
test_melt3 = melt(test3)


test_melt1$Sample = 1000
test_melt2$Sample = 5000
test_melt3$Sample = 10000

total = rbind(test_melt1, test_melt2, test_melt3)
total$Sample = as.factor(total$Sample)

colnames(total) = c("Simulation_num", "Method", "p", "Sample_size")



p1 = ggplot(total) + aes(x = Sample_size, y = -log(p), fill = Method) + geom_boxplot() + theme_classic() + labs(title="-log10(P) under different Sample size")


p1
```


Notice when sample size gets larger, we tend to have smaller p value, which is expected. Here we set effect size of AFR = EUR, therefore we lose statistical power.

&nbsp;  
&nbsp;  
&nbsp;  
&nbsp;  






# Different AFR EUR effect size

#### The following conditions remain unchanged:

num_sample = 5000  

AFR_maf = 0.05  

EUR_maf = 0.05  

num_risk_alleles = 5  


```{r, echo=FALSE}
test1 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 0, EUR_eff_size = 0)
test2 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 0, EUR_eff_size = 1)
test3 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 0, EUR_eff_size = 2)
test4 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 0)
test5 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 1)
test6 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 2)
test7 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 2, EUR_eff_size = 0)
test8 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 2, EUR_eff_size = 1)
test9 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 2, EUR_eff_size = 2)

test_melt1 = melt(test1)
test_melt2 = melt(test2)
test_melt3 = melt(test3)
test_melt4 = melt(test4)
test_melt5 = melt(test5)
test_melt6 = melt(test6)
test_melt7 = melt(test7)
test_melt8 = melt(test8)
test_melt9 = melt(test9)


test_melt1$AFR_EUR_eff = "0:0"
test_melt2$AFR_EUR_eff = "0:1"
test_melt3$AFR_EUR_eff = "0:2"
test_melt4$AFR_EUR_eff = "1:0"
test_melt5$AFR_EUR_eff = "1:1"
test_melt6$AFR_EUR_eff = "1:2"
test_melt7$AFR_EUR_eff = "2:0"
test_melt8$AFR_EUR_eff = "2:1"
test_melt9$AFR_EUR_eff = "2:2"



total = rbind(test_melt1, 
              test_melt2, 
              test_melt3,
              test_melt4, 
              test_melt5, 
              test_melt6,
              test_melt7, 
              test_melt8, 
              test_melt9
              )
total$AFR_EUR_eff = as.factor(total$AFR_EUR_eff)

colnames(total) = c("Simulation_num", "Method", "p", "AFR_EUR_eff")



p2 = ggplot(total) + aes(x = AFR_EUR_eff, y = -log10(p), fill = Method) + geom_boxplot() + theme_classic() + labs(title="-log10(P) under different AFR:EUR effect size")


p2
```


We vary the effect size of AFR and EUR, which is indicated in the x-axis (AFR:EUR). Notice when we have AFR effect size = EUR effect size, we tend to lose power. However, when they have different effect size, we tend to gain power. The most dramatic power gain is observed in AFR = 0, and EUR = 2. 


&nbsp;  
&nbsp;  
&nbsp;  
&nbsp;  



# Different causal alleles

#### The following conditions remain unchanged:

num_sample = 5000  

AFR_maf = 0.05  

EUR_maf = 0.05  

AFR_eff_size = 1  

EUR_eff_size = 1

```{r, echo=FALSE}
test1 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 1, AFR_eff_size = 1, EUR_eff_size = 1)
test2 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 1)
test3 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 20, AFR_eff_size = 1, EUR_eff_size = 1)

test_melt1 = melt(test1)
test_melt2 = melt(test2)
test_melt3 = melt(test3)


test_melt1$Causal_alleles = "1"
test_melt2$Causal_alleles = "5"
test_melt3$Causal_alleles = "20"

total = rbind(test_melt1, test_melt2, test_melt3)
total$Causal_alleles = factor(total$Causal_alleles, levels = c("1", "5", "20"))

colnames(total) = c("Simulation_num", "Method", "p", "Causal_alleles")



p3 = ggplot(total) + aes(x = Causal_alleles, y = -log(p), fill = Method) + geom_boxplot() + theme_classic() + labs(title="-log10(P) under different number of causal alleles")


p3
```



We simulate a gene with 100 variants, among 1/5/20 of them are causal variants (same direction). As expected, we gain power when there are more causal variants within a gene. We lose power here because AFR and EUR have the same effect size :(


&nbsp;  
&nbsp;  
&nbsp;  
&nbsp;  



# Different AFR/EUR minor allele frequency

#### The following conditions remain unchanged:

num_sample = 5000  

num_risk_alleles = 5

AFR_eff_size = 1  

EUR_eff_size = 1

```{r, echo=FALSE}
test1 = Retrieve_data(num_sample = 5000, AFR_maf = 0.01, EUR_maf = 0.01, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 1)
test2 = Retrieve_data(num_sample = 5000, AFR_maf = 0.01, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 1)
test3 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.01, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 1)
test4 = Retrieve_data(num_sample = 5000, AFR_maf = 0.05, EUR_maf = 0.05, num_risk_alleles = 5, AFR_eff_size = 1, EUR_eff_size = 1)

test_melt1 = melt(test1)
test_melt2 = melt(test2)
test_melt3 = melt(test3)
test_melt4 = melt(test4)


test_melt1$AFR_EUR_maf = "0.01:0.01"
test_melt2$AFR_EUR_maf = "0.01:0.05"
test_melt3$AFR_EUR_maf = "0.05:0.01"
test_melt4$AFR_EUR_maf = "0.05:0.05"

total = rbind(test_melt1, test_melt2, test_melt3, test_melt4)
total$AFR_EUR_maf = factor(total$AFR_EUR_maf, levels = c("0.01:0.01", "0.01:0.05", "0.05:0.01", "0.05:0.05"))

colnames(total) = c("Simulation_num", "Method", "p", "AFR_EUR_maf")



p4 = ggplot(total) + aes(x = AFR_EUR_maf, y = -log(p), fill = Method) + geom_boxplot() + theme_classic() + labs(title="-log10(P) under different AFR:EUR minor allele frequencies")


p4

```

In general, if we have higher minor allele frequency, we tend to have better power. 




&nbsp;  
&nbsp;  
&nbsp;  
&nbsp;  


### A summary:

Tractor burden test is meaningful when AFR and EUR causal variants have different effect size. We tend to gain power and make a more accurate estimate of the effect size. However, if their effect size is same/similar, a standard burden test is perfered.










